---
title: "High dimension reservoir"
format: gfm
editor: visual
echo: false
message: false
warning: false
prefer-html: true
---

```{r}
########## LOAD PACKAGES #####
library(dplyr)
library(parallel)
library(ggplot2)
source(file = here::here("script/FctCleanFeaturesName.R"))
set.seed(1)
```

# Setting

```{r}
df_data <- read.csv(here::here("data_obfuscated/20220117.csv"))
########## NB FEATURES #####
nb_features <- read.csv(here::here("data_obfuscated/20220117.csv")) %>%
  select(-outcome, -outcomeDate, -START_DATE) %>%
  ncol()
########## Understand gaps #####
df_data %>%
  select(P_TOUS_AGES, TESTED_TOUS_AGES, IPTCC.mean, outcome, outcomeDate, hosp) %>%
  mutate(outcomeDate = as.Date(outcomeDate)) %>%
  tidyr::pivot_longer(cols = -outcomeDate) %>%
  ggplot(mapping = aes(x = outcomeDate, y = value)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  scale_x_date(date_breaks = "months") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90),
        strip.text.y = element_text(angle = 0)) +
  labs(x = "Date")
```

The total number of features is `r nb_features`.

# PCA

```{r}
ncp = 20
pca <- df_data %>%
  select(-outcome, -outcomeDate, -START_DATE) %>%
  FactoMineR::PCA(graph = FALSE, ncp = ncp)
```

```{r}
pca$eig %>%
  as.data.frame() %>%
  tibble::rowid_to_column(var = "Nb_component") %>%
  ggplot(mapping = aes(y = Nb_component, x = `cumulative percentage of variance`)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Cumulative percentage of variance",
       y = "Nb of components")
```

```{r fig.height=10}

plotPCA <- pca$var$cos2 %>%
  as.data.frame() %>%
  tibble::rownames_to_column("features") %>%
  tidyr::pivot_longer(cols = - features) %>%
  group_by(name) %>%
  slice_max(value, n = 3) %>%
  mutate(features_clean = FctCleanFeaturesName(features),
         features_clean = forcats::fct_reorder(features_clean, value),
         name = factor(name,
                       levels = paste0("Dim.", 1:ncp),
                       labels = c("PC1: \"SAMU\"",
                                  "PC2: \"Emergency all site\"",
                                  "PC3: \"PCR+\"",
                                  "PC4: \"PCR+\"",
                                  "PC5: \"SAMU\"",
                                  "PC6: \"Weather\"",
                                  "PC7: \"Diarrhea\"",
                                  "PC8: \"PCR\"",
                                  "PC9: \"Ageusia, anosmia\"",
                                  "PC10: \"Emergency site 2\"",
                                  "PC11: \"Pediatric\"",
                                  "PC12: \"PCR+ elderly\"",
                                  "PC13: \"Dyspnea\"",
                                  "PC14: \"Fever + Dew point\"",
                                  "PC15: \"Pediatric headache\"",
                                  "PC16: \"Pediatric dyspnea\"",
                                  "PC17: \"Fever + wind speed\"",
                                  "PC18: \"Site 2 + SAMU\"",
                                  "PC19: ?",
                                  "PC20: ?"))) %>%
  ggplot(mapping = aes(y = features_clean, x = value, color = name)) +
  geom_point() +
  scale_color_viridis_d() +
  facet_wrap(name ~ ., scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(size = 9)) +
  labs(x = "Squared cosine",
       y = "")

plotPCA

ggsave(here::here("results/final_figures/plotPCA.pdf"),
       plot = plotPCA,
       width = 11,
       height = 8)
```

```{r fig.height=10}
pca$ind$coord %>%
  as.data.frame() %>%
  mutate(outcomeDate = as.Date(df_data$outcomeDate),
         outcome = df_data$outcome,
         .before = 1) %>%
  tidyr::pivot_longer(cols = -c(outcomeDate)) %>%
  mutate(name = factor(name, levels = c("outcome", paste0("Dim.", 1:ncp)))) %>%
  ggplot(mapping = aes(y = value, x = outcomeDate, color = name)) +
  geom_line() +
  scale_color_manual(values = c("black", viridis::viridis(n = ncp))) +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(name ~ ., scales = "free_y") +
  labs(x = "Date",
       y = "Value",
       color = "",
       linewidth = "")
```

# Forecast

```{r}
### 1) Load data
path_predictions <- here::here("results/final_prediction/predictions/")
ls_files <- list.files(path_predictions)
ls_files_full <- list.files(path_predictions, full.names = TRUE)
names(ls_files_full) <- gsub(ls_files, pattern = "_combined.csv", replacement = "")


df_all_temp <- lapply(ls_files_full, read.csv) %>%
  bind_rows(.id = "model") %>%
  mutate(outcomeDate = as.Date(outcomeDate),
         model = factor(model,
                        levels = c("2000_units",
                                   "GeneticSingleIs_GA",
                                   "GeneticSingleIs_GA_1000anteriorite",
                                   "GeneticSingleIs_GA_PCA",
                                   "GeneticSingleIs_RS",
                                   "SingleIs_GA",
                                   "enet_pred_RS",
                                   "xgb_pred_RS",
                                   "LSTM_PCA",
                                   "LSTM_RC",
                                   "transformers_PCA",
                                   "transformers_RC"),
                        labels = c("Reservoir FS (GA - 2000)",
                                   "Reservoir FS (GA)",
                                   "Reservoir FS (GA) all anteriority",
                                   "Reservoir FS (GA - PCA)",
                                   "Reservoir FS (RS)",
                                   "Reservoir no FS (GA)",
                                   "Elastic-net (RS)",
                                   "XGB (RS)",
                                   "LSTM-PCA",
                                   "LSTM-RC-GA",
                                   "Transformers-PCA",
                                   "Transformers-RC-GA"))) %>%
  group_by(outcomeDate, model) %>%
  filter(outcomeDate >= as.Date("2021-03-15"))

df_all <- df_all_temp %>% slice_min(hp_date) %>% mutate(update = "No") %>%
  bind_rows(df_all_temp %>% slice_max(hp_date) %>% mutate(update = "Yes")) %>%
  ungroup() %>%
  filter(!(model %in% c("Reservoir FS (GA) all anteriority",
                        "Reservoir FS (GA - 2000)",
                        "LSTM",
                        "Transformers") & update == "Yes"))
```

## Performance

```{r}
df_perf <- df_all %>%
  mutate(outcome = if_else(outcome < 10, 10, outcome),
         pred = if_else(pred < 10, 10, pred),
         hosp = if_else(hosp < 10, 10, hosp)) %>%
  group_by(outcomeDate, model, update) %>%
  summarise(outcome = unique(outcome),
            hosp = unique(hosp),
            pred = median(pred),
            .groups = "drop") %>%
  mutate(AE = abs(pred - outcome),
         RE = AE/outcome,
         baseline_AE = abs(hosp - outcome),
         AE_baseline = AE - baseline_AE,
         RE_baseline = AE/baseline_AE) %>%
  group_by(model, update) %>%
  summarise(sd_AE = sd(AE, na.rm = T),
            AE = mean(AE, na.rm = T),
            sd_AE_baseline = sd(AE_baseline, na.rm = T),
            AE_baseline = mean(AE_baseline, na.rm = T),
            sd_RE = sd(RE, na.rm = TRUE),
            RE = median(RE, na.rm = TRUE),
            sd_RE_baseline = sd(RE_baseline, na.rm = TRUE),
            RE_baseline = median(RE_baseline, na.rm = TRUE),
            .groups = "drop") %>%
  select(model, update, AE, sd_AE, AE_baseline, sd_AE_baseline, RE, sd_RE, RE_baseline, sd_RE_baseline)

df_perf %>%
  # knitr::kable(format = "latex",
  #              digits = 2)
  knitr::kable(digits = 2,
               caption = "Model performance") %>%
  kableExtra::kable_styling()

# df_all %>%
#   group_by(model) %>%
#   summarise(min = min(outcomeDate),
#             max = max(outcomeDate),
#             len = length(unique(outcomeDate)),
#             len2 = length(outcomeDate))

```

## RC aggregation

```{r}
#| label: rc-aggreg
#| fig-cap: "Aggregation of reservoir"

nboot <- 500
vec_nb_esn <- c(1:10, 20, 40)

dfreservoir <- df_all %>%
  filter(model == "Reservoir FS (GA)", update == "No") %>%
  mutate(outcome = if_else(outcome < 10, 10, outcome),
         pred = if_else(pred < 10, 10, pred),
         hosp = if_else(hosp < 10, 10, hosp))

dfRepeatReservoir <- mclapply(X = seq_len(nboot),
                              mc.cores = parallel::detectCores()-2,
                              function(boot_i){
                                lapply(X = vec_nb_esn,
                                       function(nb_esn){
                                         MAE <- dfreservoir %>%
                                           group_by(outcomeDate) %>%
                                           sample_n(nb_esn, replace = TRUE) %>%
                                           summarise(outcome = unique(outcome),
                                                     hosp = unique(hosp),
                                                     pred = median(pred),
                                                     .groups = "drop") %>%
                                           mutate(AE = abs(pred - outcome)) %>%
                                           pull(AE) %>%
                                           mean(na.rm = T)
                                         
                                         data.frame(nb_esn = nb_esn,
                                                    MAE = MAE)
                                       }) %>%
                                  bind_rows()
                              }) %>%
  bind_rows(.id = "boot_i")

plot_repeated_reservoir <- dfRepeatReservoir %>%
  mutate(MAE = if_else(MAE > 60, 60, MAE)) %>%
  group_by(nb_esn) %>%
  summarise(median_MAE = median(MAE),
            ci_inf = quantile(MAE, 0.025),
            ci_sup = quantile(MAE, 0.975)) %>%
  ggplot(mapping = aes(x = nb_esn, y = median_MAE, ymin = ci_inf, ymax = ci_sup)) +
  geom_ribbon(fill = "grey") +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(x = "Nb of Reservoir",
       y = "MAE (95% CI)")

plot_repeated_reservoir

ggsave(here::here("results/final_figures/plot_repeated_reservoir.pdf"),
       plot = plot_repeated_reservoir,
       width = 4,
       height = 4)
```

## Graphical forecast

```{r}
#| label: model-prediction
#| fig-cap: "Model prediction"

df_individual_model <- df_all %>%
  filter(update == "No") %>%
  group_by(outcomeDate, model) %>%
  summarise(outcome = unique(outcome),
            hosp = unique(hosp),
            pred = median(pred),
            .groups = "drop") %>%
  mutate(outcomeDate = as.Date(outcomeDate)) %>%
  tidyr::pivot_wider(names_from = model, values_from = pred) %>%
  tidyr::pivot_longer(cols = -c(outcomeDate)) %>%
  mutate(name = factor(name,
                       levels = c("outcome",
                                  "hosp",
                                  "Elastic-net (RS)",
                                  "XGB (RS)",
                                  "LSTM-PCA",
                                  "LSTM-RC-GA",
                                  "Transformers-RC-GA",
                                  "Transformers-PCA",
                                  "Reservoir FS (RS)",
                                  "Reservoir no FS (GA)",
                                  "Reservoir FS (GA) all anteriority",
                                  "Reservoir FS (GA - PCA)",
                                  "Reservoir FS (GA)",
                                  "Reservoir FS (GA - 2000)")),
         name = forcats::fct_recode(name,
                                    "Observed" = "outcome",
                                    "Hosp t+14" = "hosp"))

plot_figure_performance_no_update <- df_individual_model %>%
  filter(name %in% c("Hosp t+14",
                     "Observed",
                     "Reservoir FS (GA - 2000)",
                     "XGB (RS)",
                     "LSTM-PCA")) %>%
  ggplot(mapping = aes(x = outcomeDate, y = value, color = name)) +
  geom_line() +
  ggstar::geom_star(x = as.Date("2021-08-19"),
                    starshape = 23,
                    color = "white",
                    fill = "#FFB703",
                    y = 80,
                    size=5,
                    inherit.aes = FALSE) +
  annotate(geom = "text",
           label = "Holiday break",
           x = as.Date("2021-08-19"),
           y = 130,
           size=3.5,
           angle = 90) +
  scale_color_manual(values = c("black", "darkgrey", "#005F73", "#94D2BD", "#AE2012")) +
  scale_x_date(breaks = "2 month", date_labels = "%y-%m") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Date",
       y = "Hospitalizations",
       color = "",
       fill = "") +
  guides(color = guide_legend(ncol = 2))


plot_figure_performance_no_update

ggsave(here::here("results/final_figures/plot_figure_performance_no_update.pdf"),
       plot = plot_figure_performance_no_update,
       width = 4,
       height = 5)
```

```{r}
#| label: model-prediction-all
#| fig-cap: "All model prediction depending on update"

plot_figure_performance_all <- df_all %>%
  group_by(outcomeDate, model, update) %>%
  summarise(hosp = unique(hosp),
            outcome = unique(outcome),
            pred = median(pred),
            .groups = "drop") %>%
  mutate(outcomeDate = as.Date(outcomeDate),
         update = if_else(update == "Yes", "Hp updated each month", "Hp set at 2021-03-01")) %>%
  ggplot(mapping = aes(x = outcomeDate, y = pred, color = update)) +
  geom_line() +
  geom_line(mapping = aes(y = outcome, color = " Observed")) +
  geom_line(mapping = aes(y = hosp, color = "Hosp t+14")) +
  scale_color_manual(values = c("black", "darkgrey", "#AE2012", "#EE9B00")) +
  scale_x_date(breaks = "2 month", date_labels = "%y-%m") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Date",
       y = "Hospitalizations",
       color = "") +
  guides(color = guide_legend(ncol = 3)) +
  facet_wrap(model ~ ., ncol = 2)

plot_figure_performance_all

ggsave(here::here("results/final_figures/plot_figure_performance_all.pdf"),
       plot = plot_figure_performance_all,
       width = 8, height = 10)

```

# Hyperparameters

```{r}
path_hp <- here::here("results/final_prediction/hyperparameter")
ls_files_full <- list.files(path_hp, full.names = TRUE, recursive = TRUE)
names(ls_files_full) <- gsub(ls_files_full, pattern = paste0(path_hp, "/"), replacement = "")

numeric_hp <- c(
  "ridge",
  "l1_ratio",
  "spectral_radius",
  "leaking_rate",
  "input_scaling",
  "n_estimators",
  "max_depth",
  "learning_rate",
  "subsample",
  "colsample_bytree",
  "pca"
)

df_all_hp <- lapply(ls_files_full,
                    function(x) read.csv(x) %>%
                      mutate_all(as.character) %>%
                      tibble::rowid_to_column(var = "genetic_id")) %>%
  bind_rows(.id = "file_i") %>%
  mutate_at(
    .vars = c(numeric_hp, "time_seconds", "value"),
    .funs = as.numeric
  ) %>%
  tidyr::separate_wider_delim(cols = file_i,
                              delim = "/",
                              names = c("model", "date")) %>%
  mutate(model = factor(model,
                        levels = c("hp_GeneticSingleIs_GA",
                                   "hp_GeneticSingleIs_RS",
                                   "hp_SingleIs_GA",
                                   "hp_GeneticSingleIs_GA_1000anteriorite",
                                   "hp_GeneticSingleIs_GA_PCA",
                                   "hp_enet_pred_RS",
                                   "hp_xgb_pred_RS"),
                        labels = c("Reservoir FS (GA)",
                                   "Reservoir FS (RS)",
                                   "Reservoir no FS (GA)",
                                   "Reservoir FS (GA) all anteriority",
                                   "Reservoir FS (GA - PCA)",
                                   "Elastic-net (RS)",
                                   "XGB (RS)")),
         date = gsub(pattern = ".csv", x = date, replacement = ""),
         date = as.Date(date),
         date = if_else(is.na(date), as.Date("2021-03-01"), date),
         last_used_observation = date) %>%
  filter(last_used_observation < as.Date("2022-01-17"),
         value != 1000) %>%
  filter(model != "Reservoir FS (GA) all anteriority")

## save best model most important features
df_all_hp %>%
  filter(model == "Reservoir FS (GA)") %>%
  group_by(last_used_observation) %>%
  slice_min(value, n = 1) %>%
  select(c(last_used_observation, ends_with("_bin"))) %>%
  tidyr::pivot_longer(cols = -last_used_observation) %>%
  filter(value == "y") %>%
  select(last_used_observation, name) %>%
  mutate(name = gsub(name, replacement = "", pattern = "_bin")) %>%
  write.csv(here::here("results/best_features.csv"), row.names = FALSE)

## get the best 40 by date
df_all_hp_best40 <- df_all_hp %>%
  group_by(model, last_used_observation) %>%
  slice_min(value, n = 40)
```

## Numeric hyperparameters

```{r fig.height=10}
#| label: numeric-hp
#| fig-cap: "Numeric hyperparameters"

df_all_hp_best40_numeric <- df_all_hp_best40 %>%
  select(all_of(c("job_id", "model", "date", "value", numeric_hp))) %>%
  tidyr::pivot_longer(cols = numeric_hp,
                      values_to = "HP_value",
                      names_to = "HP_name") %>%
  na.omit()

ls_hpnum_plots <- df_all_hp_best40_numeric %>%
  group_by(model) %>%
  group_split() %>%
  lapply(FUN = function(df_i){
    model_title <- df_i$model %>% unique()
    df_i %>%
      group_by(last_used_observation) %>%
      mutate(rank = dense_rank(value),
             rank_factor = if_else(rank == 1, "Best", "Other")) %>%
      ungroup() %>%
      mutate(last_used_observation = last_used_observation) %>%
      arrange(desc(rank_factor)) %>%
      ggplot(mapping = aes(y = value, x = HP_value, fill = last_used_observation, size = rank_factor, color = rank_factor)) +
      geom_point(shape = 21) +
      facet_wrap(HP_name ~ ., scales = "free_x") +
      scale_fill_viridis_c(direction = -1,
                           labels = scales::date_format("%Y-%m"),
                           trans = "date",
                           breaks = as.Date(c("2021-03-01",
                                              "2021-05-01",
                                              "2021-07-01",
                                              "2021-09-01",
                                              "2021-11-01",
                                              "2022-01-01")),
                           limits = as.Date(c("2021-03-01",
                                              "2022-01-01"))) +
      scale_size_manual(values = c(4,2)) +
      scale_color_manual(values = c("red", "#AAAAAA00")) +
      scale_x_log10() +
      theme_minimal() +
      labs(x = "Hyperparameter value",
           y = "MAE",
           fill = "Update date",
           color = "",
           size = "",
           title = model_title) %>%
      return()
  })

plot_all_hp <- ggpubr::ggarrange(plotlist = ls_hpnum_plots, common.legend = TRUE, ncol = 2, nrow = 3, legend = "right")

plot_all_hp

ggsave(here::here("results/final_figures/plot_all_hp.pdf"),
       plot = plot_all_hp,
       width = 10, height = 12)
```

```{r fig.height=10}
#| label: numeric-hp-rc-ga
#| fig-cap: "Numeric hyperparameters of RC-GA"

dfGeneticSinglIS_GA_visu <- df_all_hp %>%
  filter(model == "Reservoir FS (GA)") %>%
  group_by(date) %>%
  mutate(rank = dense_rank(value),
         rank_factor = case_when(rank <= 40 ~ "Best 40",
                                 rank <= 200 ~ "Best 200",
                                 rank > 200 ~ "Other"),
         rank_factor = factor(rank_factor,
                              levels = c("Best 200",
                                         "Best 40",
                                         "Other"))) %>%
  ungroup() %>%
  select(all_of(c("rank_factor", "genetic_id", "last_used_observation", "value", "spectral_radius", "leaking_rate", "input_scaling", "ridge"))) %>%
  tidyr::pivot_longer(cols = c("spectral_radius", "leaking_rate", "input_scaling", "ridge"),
                      values_to = "HP_value", names_to = "HP_name") %>%
  mutate(value = if_else(value > 50, 50, value))

plot_genetic_algo <- ggplot(data = dfGeneticSinglIS_GA_visu,
                            mapping = aes(x = HP_value, y = value, color = genetic_id)) +
  geom_point() +
  # scale_color_brewer() +
  scale_color_viridis_c(direction = -1, option = "C", begin = 0.3) +
  labs(color = "Genetic individual") +
  ggnewscale::new_scale_color() +
  geom_point(data = dfGeneticSinglIS_GA_visu %>%
               filter(rank_factor != "Other") %>%
               arrange(rank_factor), 
             aes(color = rank_factor)) +
  scale_color_manual(values = c("black", "green")) +
  facet_grid(last_used_observation ~ HP_name, scales = "free_x") +
  scale_y_log10() +
  scale_x_log10(breaks = c(1e-10, 1e-5, 1, 1e5),
                labels = c("-10", "-5", "1", "5")) +
  theme_minimal() +
  theme(strip.text.y = element_text(angle = 0),
        legend.position = "bottom") +
  labs(x = "Log(Hyperparameter value)",
       y = "MAE",
       color = "")

# dfGeneticSinglIS_GA_visu %>%
#   filter(last_used_observation == "2022-01-01", rank_factor == "Best 40") %>%
#   group_by(HP_name) %>%
#   summarise(min(HP_value), max(HP_value))

plot_genetic_algo

ggsave(here::here("results/final_figures/plot_genetic_algo.pdf"),
       plot = plot_genetic_algo,
       useDingbats = TRUE,
       width = 5,
       height = 7)
```

## Categorical hyperparameters

```{r}
#| label: feature-selection-rc-ga
#| fig-cap: "Feature selection by RC-GA"

categorical_features <- grep(colnames(df_all_hp_best40), pattern = "_bin$", value = TRUE)
df_all_hp_best40_categorical <- df_all_hp_best40 %>%
  select(all_of(c("job_id", "model", "last_used_observation", "value", categorical_features))) %>%
  tidyr::pivot_longer(cols = categorical_features,
                      values_to = "HP_value",
                      names_to = "HP_name") %>%
  na.omit() %>%
  group_by(model, last_used_observation, HP_name) %>%
  summarise(freq_select = mean(HP_value == "y"),
            .groups = "drop")

ls_hpcat_plots <- list()
df_all_hp_best40_categorical %>%
  mutate(HP_name = FctCleanFeaturesName(HP_name)) %>%
  group_by(model) %>%
  group_split() %>%
  lapply(FUN = function(df_i){
    
    model_title <- df_i$model %>% unique() %>% as.character
    
    vec_features_i <- df_i %>%
      filter(last_used_observation == min(last_used_observation)) %>%
      slice_max(freq_select, n = 10) %>%
      pull(HP_name)
    
    plot_importance <- df_i %>%
      filter(HP_name %in% vec_features_i) %>%
      ggplot(mapping = aes(x = last_used_observation, color = freq_select, y = HP_name,
                           linewidth = freq_select)) +
      geom_line() +
      scale_color_distiller(breaks = seq(0,1,by=0.5),
                            limits = c(0,1),
                            direction = 1,
                            palette = "Greys") +
      scale_linewidth_continuous(breaks = seq(0,1,by=0.5),
                                 limits = c(0,1)) +
      scale_x_date(date_labels = "%y-%m") +
      theme_minimal() +
      labs(x = "Update date",
           y = "",
           color = "Freq. select",
           linewidth = "Freq. select",
           title = model_title) +
      theme(legend.position = "bottom",
            legend.box="vertical")
    
    ls_hpcat_plots[[model_title]] <<- plot_importance
    return()
  })

##### elastic-net and xgb importance when updating hyperparameter monthly for comparison
list_importance_enet_xgb <- list("Elastic-net" = here::here("results/final_prediction/importance/enet_pred_RS_importance_combined.csv"),
                                 "XGBoost" = here::here("results/final_prediction/importance/xgb_pred_RS_importance_combined.csv"))

lapply(names(list_importance_enet_xgb),
       function(x){
         df_imp_enet <- data.table::fread(list_importance_enet_xgb[[x]]) %>%
           mutate(last_used_observation = hp_date + 14) %>%
           group_by(outcomeDate) %>%
           slice_max(last_used_observation) %>%
           group_by(last_used_observation, features) %>%
           summarise(importance = mean(importance),
                     .groups = "drop")
         
         best_enet_features_first_date <- df_imp_enet %>%
           slice_min(last_used_observation) %>%
           slice_max(abs(importance), n = 10) %>%
           pull(features)
         
         color_lab <- ifelse(x == "XGBoost", "Mean Gain", "Mean Beta")
         
         plot_importance <- df_imp_enet %>%
           filter(features %in% best_enet_features_first_date) %>%
           mutate(features = FctCleanFeaturesName(features),
                  features = factor(features),
                  features = forcats::fct_reorder(features, abs(importance))) %>%
           ggplot(mapping = aes(x = last_used_observation, color = importance, y = features,
                                linewidth = importance)) +
           geom_line() +
           scale_color_distiller(direction = 1,
                                 palette = "Greys") +
           scale_x_date(date_labels = "%y-%m") +
           theme_minimal() +
           labs(x = "Update date",
                y = "",
                color = color_lab,
                linewidth = color_lab,
                title = x) +
           theme(legend.position = "bottom",
                 legend.box="vertical")
         
         ls_hpcat_plots[[x]] <<- plot_importance
         return()
       })

ls_hpcat_plots$`Reservoir FS (GA)`

ggsave(here::here("results/final_figures/plot_feature_imp_RCGA.pdf"),
       plot = ls_hpcat_plots$`Reservoir FS (GA)`,
       width = 6, height = 5)

```

```{r fig.height=15}
#| label: feature-selection-all
#| fig-cap: "Feature selection"

plot_feature_imp_temp <- ggpubr::ggarrange(plotlist = ls_hpcat_plots[1:2],
                                           ncol = 2, nrow = 1,
                                           legend = "bottom",
                                           common.legend = TRUE)

plot_feature_imp <- gridExtra::grid.arrange(plot_feature_imp_temp,
                                            ls_hpcat_plots$`Elastic-net`,
                                            ls_hpcat_plots$XGBoost,
                                            layout_matrix = rbind(c(1, 1),
                                                                  c(2, 3)))


ggpubr::ggarrange(plotlist = ls_hpcat_plots,
                  ncol = 1, nrow = 4,
                  legend = "bottom")

ggsave(here::here("results/final_figures/plot_feature_imp.pdf"),
       plot = plot_feature_imp,
       width = 12, height = 9)

```

## Ridge update of 2000 units reservoir

```{r}
#| label: ridge-for-2000-units
#| fig-cap: "Ridge hyperparameter of 2000 units RC-GA"

list.files(path = here::here("results/2000units/"), full.names = T) %>%
  lapply(data.table::fread) %>%
  bind_rows() %>%
  ggplot(mapping = aes(x = ridge, y = value)) +
  geom_point() +
  scale_x_log10(breaks = 10^(2:7)) +
  scale_y_log10() +
  theme_minimal() +
  labs(x = "ridge", y = "MAE")
```

# Failure of monthly update

```{r}
############# explore why hp update did not work for GA
df_leaking_rate_RCGA <- df_all_hp %>%
  filter(model == "Reservoir FS (GA)") %>%
  mutate(hp_date = date+14) %>%
  select(job_id, leaking_rate, hp_date) %>% 
  distinct()

df_by_lr <- df_all %>%
  filter(model == "Reservoir FS (GA)", update == "No") %>%
  mutate(job_id = gsub(x = trial, pattern = "^trial_|_train365$", replacement = ""),
         hp_date = as.Date(hp_date)) %>%
  left_join(df_leaking_rate_RCGA, by = c("job_id", "hp_date")) %>%
  mutate(leaking_rate = factor(leaking_rate < 1e-3,
                               levels = c(T, F),
                               labels = c("< 1e-3", "> 1e-3")),
         outcomeDate = as.Date(outcomeDate))

df_perf_by_lr <- df_by_lr %>%
  mutate(outcome = if_else(outcome < 10, 10, outcome),
         pred = if_else(pred < 10, 10, pred),
         hosp = if_else(hosp < 10, 10, hosp)) %>%
  group_by(outcomeDate, model, leaking_rate, update) %>%
  summarise(outcome = unique(outcome),
            hosp = unique(hosp),
            pred = median(pred),
            n = n(),
            .groups = "drop") %>%
  mutate(AE = abs(pred - outcome),
         RE = AE/outcome,
         baseline_AE = abs(hosp - outcome),
         AE_baseline = AE - baseline_AE,
         RE_baseline = AE/baseline_AE) %>%
  group_by(model, update, leaking_rate, n) %>%
  summarise(AE = mean(AE, na.rm = T),
            AE_baseline = mean(AE_baseline, na.rm = T),
            RE = median(RE, na.rm = TRUE),
            RE_baseline = median(RE_baseline, na.rm = TRUE),
            .groups = "drop")

df_perf_by_lr %>%
  # knitr::kable(format = "latex", booktabs = TRUE, digits = 2) %>%
  knitr::kable(caption = "Performance by leaking rate") %>%
  kableExtra::kable_styling()
```

```{r}
#| label: prediction-by-lr
#| fig-cap: "Prediction by leaking rate"

plot_RCGA_noupdate_by_lr <- df_by_lr %>%
  select(outcomeDate, outcome, hosp, pred, leaking_rate) %>%
  tidyr::pivot_wider(values_from = pred, names_from = leaking_rate, values_fn = median, names_prefix = "leaking rate ") %>%
  tidyr::pivot_longer(cols = -outcomeDate) %>%
  mutate(name = factor(name,
                       levels = c("outcome",
                                  "hosp",
                                  "leaking rate < 1e-3",
                                  "leaking rate > 1e-3"),
                       labels = c("Observed",
                                  "Hosp t+14",
                                  "leaking rate < 1e-3",
                                  "leaking rate > 1e-3"))) %>%
  ggplot(mapping = aes(x = outcomeDate, y = value, color = name)) +
  geom_line() +
  scale_color_manual(values = c("black", "darkgrey", "#FB8500", "#219EBC")) +
  scale_x_date(breaks = "2 month", date_labels = "%y-%m") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Date",
       y = "Hospitalizations",
       color = "") +
  guides(color = guide_legend(ncol = 2))

plot_RCGA_noupdate_by_lr

ggsave(here::here("results/final_figures/plot_RCGA_noupdate_by_lr.pdf"),
       plot = plot_RCGA_noupdate_by_lr,
       height = 4,
       width = 8)
```

# Computing time

<!-- ```{r} -->

<!-- #| eval: false -->

<!-- ### redo on local machine -->

<!-- df_time_hp <- df_all_hp_best40 %>% -->

<!--   slice_min(date) %>% -->

<!--   filter(model %in% c("Reservoir FS (RS)", -->

<!--                       "Elastic-net (RS)", -->

<!--                       "XGB (RS)")) %>% -->

<!--   mutate(nb_features = select(., ends_with("_bin")) %>% mutate_all(.funs = function(x) x == "y") %>% rowSums()) %>% -->

<!--   select(model, time_seconds, all_of(numeric_hp), nb_features) %>% -->

<!--   tidyr::pivot_longer(cols = -c("model", "time_seconds")) %>% -->

<!--   na.omit() -->

<!-- read.csv("data_obfuscated/20210329.csv") %>% -->

<!--   select(-outcome, -outcomeDate, -START_DATE) %>% -->

<!--   dim() -->

<!-- plot_time_hp <- df_time_hp %>% -->

<!--   ggplot(mapping = aes(x = value, y = time_seconds)) + -->

<!--   geom_point(size = 0.01) + -->

<!--   scale_x_log10() + -->

<!--   facet_wrap(model ~ name, scales = "free", ncol = 3) + -->

<!--   theme_minimal() -->

<!-- df_time_hp %>% -->

<!--   filter(name == "nb_features") %>% -->

<!--   reframe(quantile(value, probs = c(0.25, 0.5, 0.75))) -->

<!-- ggsave(here::here("results/final_figures/plot_time_hp.pdf"), -->

<!--        plot = plot_time_hp, -->

<!--        useDingbats = TRUE, -->

<!--        height = 7, -->

<!--        width = 7) -->

<!-- df_all_hp %>% -->

<!--   slice_min(date) %>% -->

<!--   group_by(model) %>% -->

<!--   summarise(q1 = quantile(time_seconds, 0.25), -->

<!--             median = quantile(time_seconds, 0.5), -->

<!--             q3 = quantile(time_seconds, 0.75)) -->

<!-- ``` -->

We evaluated the time needed to train and forecast on January 2022
month.

```{r}
c("results/timing.csv",
  "results/timing_LSTM.csv",
  "results/timing_TRASNFORMERS_PCA.csv") %>%
  lapply(function(file_i){
    read.csv(here::here(file_i))
  }) %>%
  bind_rows() %>%
  select(Scenario, "Time (s)" = Elapsed.Time) %>%
  mutate(Scenario = factor(Scenario,
                           levels = c("enet_pred_RS",
                                      "xgb_pred_RS",
                                      "GeneticSingleIs_GA",
                                      "GeneticSingleIs_GA_1000",
                                      "LSTM",
                                      "TRANSFORMER"),
                           labels = c("Elastic-net (RS)",
                                      "XGB (RS)",
                                      "Reservoir FS (GA)",
                                      "Reservoir FS (GA - 2000)",
                                      "LSTM",
                                      "Transformers"))) %>%
  knitr::kable(caption = "Time on january month by algorithm. For RCGA, 40 RC where performed.") %>%
  kableExtra::kable_styling()
```

# Reservoir vs raw features importance

```{r}
#| label: reservoir-vs-input
#| fig-cap: "Reservoir layer vs input layer importance according ot output layer"

df_imp_reservoir_temp <- data.table::fread(here::here("results/final_prediction/importance/GeneticSingleIs_GA_importance_combined.csv")) %>%
  slice_min(hp_date)

vec_quantile <- c(0.25, .75)
quantile_function <- lapply(vec_quantile,
                            function(quantile) function(x) quantile(abs(x), quantile))
names(quantile_function) <- paste0("q", vec_quantile*100)

plot_coef_input_vs_reservoir <- df_imp_reservoir_temp %>%
  ungroup() %>%
  select(features, importance, outcomeDate) %>%
  mutate(reservoir_features = factor(grepl("reservoir", x = features),
                                     levels = c(T, F),
                                     labels = c("Reservoir",
                                                "Input layer"))) %>%
  group_by(outcomeDate, reservoir_features) %>%
  summarise_at(.vars = "importance",
               .funs = quantile_function) %>%
  ggplot(mapping = aes(x = outcomeDate, fill = reservoir_features, ymin = q25, ymax = q75)) +
  geom_ribbon(alpha = 0.7) +
  theme_minimal() +
  scale_fill_manual(values = c("#FB8500", "#219EBC")) +
  scale_y_log10() +
  scale_x_date(breaks = "2 month", date_labels = "%y-%m") +
  labs(fill = "IQR of output layer coefficients",
       x = "Date",
       y = "Output layer coefficients") +
  theme(legend.position = "bottom")

plot_reservoir_vs_input_importance <- ggpubr::ggarrange(plot_figure_performance_no_update + ggplot2::ggtitle("A"),
                                                        plot_coef_input_vs_reservoir + ggplot2::ggtitle("B"),
                                                        nrow = 2, ncol = 1,
                                                        heights = c(0.65, 0.35))

ggsave(here::here("results/final_figures/plot_reservoir_vs_input_importance.pdf"),
       plot = plot_reservoir_vs_input_importance,
       width = 5, height = 6)

```

# Number of parameters

```{r}
ls_importance <- list(RCGA = here::here("results/final_prediction/importance/GeneticSingleIs_GA_importance_combined.csv"),
                      RCGAPCA = here::here("results/final_prediction/importance/GeneticSingleIs_GA_PCA_importance_combined.csv"),
                      Enet = here::here("results/final_prediction/importance/enet_pred_RS_importance_combined.csv"),
                      XGB = here::here("results/final_prediction/importance/xgb_pred_RS_importance_combined.csv"))

df_nb_hp <- lapply(ls_importance,
                   FUN = function(path_i){
                     path_i %>%
                       data.table::fread() %>%
                       slice_max(hp_date) %>%
                       slice_max(outcomeDate) %>%
                       select(nb_param) %>%
                       reframe(min(nb_param), max(nb_param))
                   }) %>%
  bind_rows(.id = "model")

df_nb_hp %>%
  # knitr::kable(format = "latex", booktabs = TRUE, digits = 2) %>%
  knitr::kable(caption = "Number of parameters") %>%
  kableExtra::kable_styling()
```

```{r}
df_nb_features <- lapply(names(ls_importance),
                         FUN = function(name_path_i){
                           path_i <- ls_importance[[name_path_i]]
                           df_i <- path_i %>%
                             data.table::fread() %>%
                             slice_min(hp_date)
                           
                           if(name_path_i == "Enet"){
                             df_i2 <- df_i %>%
                               filter(importance != 0)
                           }
                           if(name_path_i == "XGB"){
                             df_i2 <- df_i
                           }
                           if(name_path_i %in% c("RCGA", "RCGAPCA")){
                             df_i2 <- df_i %>%
                               filter(!grepl(features, pattern = "reservoir"))
                           }
                           df_i2 %>%
                             group_by(outcomeDate, trial) %>%
                             summarise(nb_features = n(), .groups = "drop") %>%
                             summarise(q25 = quantile(nb_features, .25),
                                       q50 = quantile(nb_features, .5),
                                       q75 = quantile(nb_features, .75)) %>%
                             mutate(model = name_path_i, .before = 1) %>%
                             return(.)
                           
                         }) %>%
  bind_rows()

df_nb_features %>%
  knitr::kable(caption = "Number of features") %>%
  kableExtra::kable_styling()
```

# Performance by date update

## Confounding matrix

```{r}
#| label: performance-by-date-update
#| fig-cap: "Performance by date update"

date_breaks_period <- df_all_temp %>%
  group_by(hp_date) %>%
  summarise(min = min(outcomeDate)) %>%
  pull(min) %>%
  unique() %>%
  as.Date()

plot_perf_by_period <- df_all_temp %>%
  ungroup() %>%
  mutate(outcomeDate = as.Date(outcomeDate),
         periods = cut.Date(x = outcomeDate, breaks = c(date_breaks_period, Inf)),
         outcome = if_else(outcome < 10, 10, outcome),
         pred = if_else(pred < 10, 10, pred),
         hosp = if_else(hosp < 10, 10, hosp)) %>%
  group_by(outcomeDate, model, hp_date, periods) %>%
  summarise(outcome = unique(outcome),
            hosp = unique(hosp),
            pred = median(pred),
            .groups = "drop") %>%
  mutate(AE = abs(pred - outcome),
         RE = AE/outcome,
         baseline_AE = abs(hosp - outcome),
         AE_baseline = AE - baseline_AE,
         RE_baseline = AE/baseline_AE) %>%
  group_by(model, hp_date, periods) %>%
  summarise(MAE = mean(AE, na.rm = TRUE)) %>%
  group_by(model, periods) %>%
  mutate(scaled_MAE = scale(MAE),
         scaled_MAE = if_else(is.na(scaled_MAE), 0, scaled_MAE)) %>%
  filter(model %in% c("Reservoir FS (GA)")) %>%
  mutate(hp_date = gsub(hp_date, pattern = "-15$", replacement = "-01"),
         hp_date = gsub(hp_date, pattern = "-16$", replacement = "-02")) %>%
  ggplot(mapping = aes(x = periods, y = hp_date, fill = scaled_MAE)) +
  geom_tile() +
  scale_fill_distiller(palette = "RdBu") +
  facet_grid(model ~ .) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Periods",
       y = "Hyperparameter update date",
       fill = "Scaled MAE")

plot_perf_by_period

ggsave(here::here("results/final_figures/plot_perf_by_period.pdf"),
       plot = plot_perf_by_period,
       width = 5, height = 4)
```

## Towards more agile update

```{r}
df_all_temp_daily_update <- df_all_temp %>%
  filter(model == "Reservoir FS (GA)") %>%
  mutate(outcomeDate = as.Date(outcomeDate),
         outcome_bis = if_else(outcome < 10, 10, outcome),
         pred_bis = if_else(pred < 10, 10, pred),
         hosp_bis = if_else(hosp < 10, 10, hosp),
         AE = abs(pred_bis - outcome_bis)) %>%
  select(model, trial, hp_date, pred, outcome, hosp, AE, outcomeDate)

vecAllDates <- as.Date(unique(df_all_temp_daily_update$outcomeDate))
vec_anteriority <- c(0, 1, 3, 7, 14, 28, Inf)
combinations <- expand.grid(anteriority = vec_anteriority, Date = vecAllDates)
result_list <- split(combinations, 1:nrow(combinations))
forecast_delay = 14

df_res <- mclapply(result_list,
                   mc.cores = detectCores()-2,
                   function(vec_i){
                     date_i_pred <- as.Date(vec_i$Date)
                     anteriority_i <- vec_i$anteriority
                     
                     date_i_start <- date_i_pred-14
                     
                     nb_model_avail <- df_all_temp_daily_update %>%
                       filter(outcomeDate == date_i_pred) %>%
                       nrow()
                     
                     if(nb_model_avail == 40){
                       best_previous2weeks <- df_all_temp_daily_update %>%
                         filter(outcomeDate == date_i_pred) %>%
                         ungroup()
                       
                     } else {
                       best_previous2weeks <- df_all_temp_daily_update %>%
                         filter(outcomeDate <= date_i_start,
                                outcomeDate >= date_i_start-anteriority_i) %>%
                         group_by(model, trial, hp_date) %>%
                         summarise(MAE = mean(AE), .groups = "drop") %>%
                         group_by(model) %>%
                         slice_min(MAE, n = 40) %>%
                         ungroup()
                     }
                     
                     res <- best_previous2weeks %>%
                       mutate(Anteriority = anteriority_i) %>%
                       select(Anteriority, model, trial, hp_date) %>%
                       left_join(df_all_temp_daily_update, by = c("model", "trial", "hp_date")) %>%
                       filter(outcomeDate == (date_i_pred)) %>%
                       return(res)
                   }) %>%
  bind_rows()

```

```{r}
df_res %>%
  group_by(outcomeDate, Anteriority) %>%
  summarise(pred = median(pred),
            outcome = unique(outcome),
            .groups = "drop") %>%
  mutate(outcome_bis = if_else(outcome < 10, 10, outcome),
         pred_bis = if_else(pred < 10, 10, pred),
         AE = abs(pred_bis - outcome_bis)) %>%
  group_by(Anteriority) %>%
  summarise(MAE = mean(AE, na.rm = TRUE)) %>%
  knitr::kable(caption = "Performance while selecting best hp sets among all of the best hp sets of very month. Results depending on the evaluation period") %>%
  kableExtra::kable_styling()
```

```{r}
#| label: performance-agile-update
#| fig-cap: "Forecast based on agile hp update where best hyperparameters from every 40 best reservoir depending on the evaluation period"

df_pred_by_anteriority <- df_res %>%
  group_by(outcomeDate, Anteriority) %>%
  summarise(outcome = unique(outcome),
            hosp = unique(hosp),
            pred = median(pred),
            .groups = "drop") %>%
  mutate(outcomeDate = as.Date(outcomeDate)) %>%
  tidyr::pivot_wider(names_from = Anteriority, values_from = pred) %>%
  tidyr::pivot_longer(cols = -c(outcomeDate)) %>%
  bind_rows(df_individual_model %>% filter(name == "Reservoir FS (GA)")) %>%
  mutate(name = factor(name,
                       levels = c("outcome",
                                  "hosp",
                                  "Reservoir FS (GA)",
                                  vec_anteriority)),
         name = forcats::fct_recode(name,
                                    "Observed" = "outcome",
                                    "Hosp t+14" = "hosp"))

plot_anteriority <- df_pred_by_anteriority %>%
  ggplot(mapping = aes(x = outcomeDate, y = value, color = name)) +
  geom_line() +
  scale_color_manual(values = c("black", "darkgrey", "#AE2012", viridis::mako(n = length(vec_anteriority)))) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(x = "Date",
       y = "Hospitalizations",
       color = "") +
  guides(color = guide_legend(ncol = 2))

plot_anteriority
```
